---
title: "Reproducible Research"
author: "Haralampi Bageski"
date: "Friday, October 18, 2014"
output: html_document
---

##Loading and preprocessing the data...
```{r read_table,echo=TRUE}
<<<<<<< HEAD
#setting of current directory differs, you should set the path to your directory, hence do not follow the next line of code
setwd("D:/HARI/university subjects/COMPUTER SCIENCE/Fun staf/Reproducible Research/Project 1")
library(utils)
=======
library(utils)
unzip("activity.zip")
>>>>>>> e6305e66cc6af87889714682eb1acd36c0bc77bd
activity<-read.csv("activity.csv")
```

  
##What is mean total number of steps taken per day?
```{r histogram,echo=TRUE,fig.height=4}
library(stats)
filter <-aggregate(steps~date ,activity,FUN=sum) #creates a datatable with columns date    and total-steps-per-day
num_of_days <- length(filter$date)
<<<<<<< HEAD
hist(filter$steps,num_of_days)  #length(filter$date) = 53 : the number of days
=======
hist(filter$steps,num_of_days) #length(filter$date) = 53 : the number of days
>>>>>>> e6305e66cc6af87889714682eb1acd36c0bc77bd
mean_step <- mean(filter$steps)
mean_step
median_step <- median(filter$steps)
median_step
```

##What is the average daily activity pattern?
```{r time_series,echo=TRUE,fig.height=4}
day_pattern <- activity[,c(1,3)]
plot(day_pattern,main="Basic plot with no modifications")

day_pattern <- aggregate(steps ~ interval, day_pattern, mean)
plot(day_pattern,type = "l",ylab="avg steps per interval",main="Activeness per time period of the day")
most_steps <- max(aggregate(steps ~ interval, day_pattern, mean)[,2],na.rm=T)
most_steps

which(day_pattern$steps == most_steps, arr.ind = TRUE)  #gives the index
active_interval <- (day_pattern[104,1])      #since intervals start from 0
active_interval
```


##Imputing missing values...
```{r missing_data,echo=TRUE,fig.height=4}
filled_activity <- activity
Steps <- filled_activity$steps
NArows <- Steps[is.na(Steps)]
length(NArows)
ndices_of_NAs <- which(is.na(Steps))
filled_activity[ndices_of_NAs,1] <-day_pattern[filled_activity[ndices_of_NAs,3]/5+1,2]  #On the right hand side activity[indices_of_NAs,3] gives the interval we want to use, but to get the row of that interval, we must apply /5+1
```


##Finding meand and median on the new data set...
```{r filled_data,echo=TRUE,fig.height=4}
knitr::opts_chunk$set(cache=TRUE)       #it makes the variables accessible from other                 code chunks
filter2 <-aggregate(steps~date ,filled_activity,FUN=sum) #creates a datatable with   columns date    and total-steps-per-day
num_of_days2 <- length(filter2$date)
hist(filter2$steps,num_of_days2)      #length(filter$date) = 53 : the number of days
mean(filter2$steps)
median(filter2$steps)
#day_pattern2 <- aggregate(steps ~ interval, filter2, mean)
#plot(day_pattern2,type = "l",ylab="avg steps per interval",main="Activeness per time period of the day")
```
  We can conclude that the new mean and median differs from the old mean `r mean_step` and old median `r median_step`. By my calculations we get that the mean and the median should differ more that what we get before filling empty data. The reason for that is the negligence of outliers.
  
  
##Are there differences in activity patterns between weekdays and weekends?
```{r weekdays,echo=TRUE,fig.height=4}
df = data.frame(date=filled_activity$date)
wkday <-c("Monday","Tuesday","Wednesday","Thursday","Friday")
df$day <- weekdays(as.Date(df$date))
filled_activity$weekday <-df$day %in% wkday  #1 if weekday, 0 otherwise
head(filled_activity)

library(sqldf) 
weekend_pattern <- sqldf("select interval,avg(steps),weekday as avg_steps from filled_activity where weekday='0' group by interval,weekday")

<<<<<<< HEAD
png("part3.png", width=480, height=480)
plot(weekend_pattern[,-3],type = "l",ylab="avg steps per interval for weekends",main="Activeness in weekdends with black, in weekdays with blue")

weekday_pattern <- sqldf("select interval,avg(steps),weekday as avg_steps from filled_activity where weekday='1' group by interval,weekday")

lines(weekday_pattern[,-3],type = "l",ylab="avg steps per interval for weekdays",main="Activeness in weekdays",col="blue")
=======
plot(weekend_pattern[,-3],type = "l",ylab="avg steps per interval for weekends",main="Activeness in weekdends")

weekday_pattern <- sqldf("select interval,avg(steps),weekday as avg_steps from filled_activity where weekday='1' group by interval,weekday")

plot(weekday_pattern[,-3],type = "l",ylab="avg steps per interval for weekdays",main="Activeness in weekdays")
>>>>>>> e6305e66cc6af87889714682eb1acd36c0bc77bd
```
  We can see from the two above graphs that people start later in the morning during weekends (interval 600-900) but in overall do more steps.
